Variance reduction for MCMC methods via martingale representations

### 2020-03-21 aap/Revision/vr-mcmc-revision-new-new.tex
update

where $N\in\mathbb N_0$ is the length of the burn-in period and $n\in\mathbb N$

@@

for any measurable function $g.$  Hence the r.v.  \(M_{K,n}^N(f)\) has zero mean and can be viewed as a control variate.

@@

A_{q,k}(y)=\sum_{r=0}^q \bar{a}_{r,k}(y),\quad q\in \mathbb{N}.

@@

In   Section~\ref{sec:ula_analysis}   we provide a detailed  theoretical  analysis of  $\mathsf{Var}[\pi_{K,n}^N(f)]$  for the  ULA  algorithm (see Example~\ref{exam:langevin-algorithm}).   

@@

$A_{q,k}$ for $q=N+1,\ldots,N+n,$ $k=1,\ldots,K. $ We first generate \(\NtrainPath\)  paths  of the chain \(X\)

@@

\TrainSet=\Bigl\{(X^{(s)}_{N+1},\ldots,X^{(s)}_{N+n}),\quad s=1,\ldots,\NtrainPath\Bigr\}

@@

where \(\Psi\) is a  class of  real-valued functions on \(\mathbb{R}^d\) and $\xi^{(s)}_{N+1},\ldots,\xi^{(s)}_{N+n}$ are iid random variables with distribution $P_\xi$ used to construct the $s$-th trajectory of $X.$   As usual in regression analysis (see Section~11 in \cite{gyorfi2006distribution})  we  truncate  the least squares estimates $(\widetilde{A}_{q,k})$ and set  

@@

\widetilde{A}_{q,k}(y), & |\widetilde{A}_{q,k}(y)|\leq U_k,

@@

U_k \,\mathrm{sign}(\widetilde{A}_{q,k}(y)), & \mathrm{otherwise}. 

@@

 for a sequence of positive real numbers $U_k,$ $k=1,\ldots,K.$

@@

\PE[A^2_{l,k}(X^x_{N+l-1})]\leq e^{\varkappa n\gamma^2}R^2_k ,\quad k=1,2,\ldots, \quad l=1,\ldots,n,

@@

 Note that the above estimate is rather rough as we ignore here the fact that the difference between $\widehat A_{n-l,k}$ and $A_{n-l,k}$ is small.
If the class of functions $\Psi$ is a linear one of dimension $\mathtt{D}_\Psi,$ then the cost of computing $\widehat \pi_{K,n}^N(f)$ is of order $ \mathtt{D}_\Psi K T^2 n$. At the same time the variance of the standard estimate $ \pi_{n}^N(f)$ is of order $1/(n\gamma)$ and this bound can not be improved, see Remark~\ref{rem:var-low}. Thus, the ratio of the corresponding cost-variances is of order

@@

\begin{lem} 
\label{lem:variance}
Let $f$ be a bounded Borel function and $(X^x_{p})_{p \in \nset_0}$ be a Markov chain generated by ULA with constant step size $\gamma$, satisfying {\bf (H1)} and {\bf (H2)}. Assume also that $n\gamma > C_0$ for some fixed $C_0,$ then it holds

@@

 Set $V(x) = 1 + \|x\|^2,$ then  it holds for $q=1,\ldots,n,$ and all $x$

@@

\right)\leq C_{K}\gamma^{-1}e^{\varkappa n\gamma^2}(1+V(x)),\quad q=1,\ldots,n,

>>>>

where $N\in\mathbb N_0$ is the length of the burn-in period and $n\in\mathbb N$ is

@@

for any integrable function $g.$  Hence the r.v.  \(M_{K,n}^N(f)\) has zero mean and can be viewed as a control variate.

@@

A_{q,k}(y)=\sum_{r=0}^q \bar{a}_{r,k}(y),\quad q=0,\ldots,n-1.

@@

In   Section~\ref{sec:ula_analysis}   we provide a detailed  theoretical  analysis of  this decay  for   ULA   (see Example~\ref{exam:langevin-algorithm}).   

@@

$A_{q,k}$ for $q=0,\ldots,n-1,$ $k=1,\ldots,K. $ We first generate \(\NtrainPath\)  paths  of the chain \(X\)

@@

\TrainSet=\Bigl\{(X^{(s)}_{1},\ldots,X^{(s)}_{N+n}),\quad s=1,\ldots,\NtrainPath\Bigr\}

@@

where \(\Psi\) is a  class of  real-valued functions on \(\mathbb{R}^d\) and $\xi^{(s)}_{1},\ldots,\xi^{(s)}_{N+n}$ are iid random variables with distribution $P_\xi$ used to construct the $s$-th trajectory of $X.$   As usual in regression analysis (see Section~11 in \cite{gyorfi2006distribution}),  we  truncate  the least squares estimates $(\widetilde{A}_{q,k})$ and set  

@@

\widetilde{A}_{q,k}(y), & |\widetilde{A}_{q,k}(y)|\leq W_k,

@@

W_k \,\mathrm{sign}(\widetilde{A}_{q,k}(y)), & \mathrm{otherwise}. 

@@

 for a sequence of positive real numbers $W_k,$ $k=1,\ldots,K.$

@@

\PE[A^2_{l,k}(X^x_{N+l-1})]\leq e^{\varkappa n\gamma^2}R^2_k ,\quad k=1,2,\ldots, \quad l=0,\ldots,n-1,

@@

 Note that the above estimate is rather rough, as we ignore here the fact that the difference between $\widehat A_{n-l,k}$ and $A_{n-l,k}$ is small.
If the class of functions $\Psi$ is a linear one of dimension $\mathtt{D}_\Psi,$ then the cost of computing $\widehat \pi_{K,n}^N(f)$ is of order $ \mathtt{D}_\Psi K T^2 n$. At the same time the variance of the standard estimate $ \pi_{n}^N(f)$ is of order $1/(n\gamma)$ and this bound can not be improved, see Lemma~\ref{lem:variance} and Remark~\ref{rem:var-low}. Thus, the ratio of the corresponding cost-variances is of order

@@

\begin{lem} 
\label{lem:variance}
Let $f$ be a bounded Borel function and $(X^x_{p})_{p \in \nset_0}$ be a Markov chain generated by ULA with constant step size $\gamma$, satisfying {\bf (H1)} and {\bf (H2)}. Assume also that $n\gamma > c_0$ for some fixed $c_0>0,$ then it holds

@@

 Set $V(x) = 1 + \|x\|^2,$ then  it holds for $q=0,\ldots,n-1,$ and all $x$

@@

\right)\leq C_{K}\gamma^{-1}e^{\varkappa n\gamma^2}(1+V(x)),\quad q=0,\ldots,n-1,
